---
title: "Homework 2"
author: "Lily Wang"
date: "9/22/2020"
output: github_document
---

This is my solution to Homework 2.

```{r libraries}
library(tidyverse)
library(readxl)
```

## Problem 1

Read in and clean the Mr. Trash Wheel dataset:
```{r clean_trashwheel}
trashwheel_df <- 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
```

Read in and clean the 2017 and 2018 precipitation data:
```{r clean_precip}
precip2017_df <-
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

precip2018_df <-
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

Make a dataframe with month names. Combine the two precipitation datasets and join the month dataframe with the result:

```{r change_month}
month_df <-
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df <-
  bind_rows(precip2017_df, precip2018_df) %>% 
  left_join(., month_df, by = "month") %>% 
  select(year, month_name, total) %>% 
  rename(month = month_name)
```
The Trash Wheel dataset contains information from the Mr. Trash Wheel trash collector located in Baltimore, MD. The dataset provides information on year, month, and amount of trash collected, including specific types of trash. There are a total of `r nrow(trashwheel_df)` rows and `r ncol(trashwheel_df)` columns in the final dataset. For available data, the total precipitation in 2018 was `r sum(pull(precip2018_df, total))` in. The median number of sports balls collected in 2017 was `r median(pull(precip2017_df, total))`.


## Problem 2

Read in and clean the NYC Transit dataset:

```{r clean_nyctransit}
nyctransit_df <-
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>% 
  mutate(entry = if_else(entry == "YES", TRUE, FALSE))
```

The NYC Transit dataset contains information such as location, routes served, entrance type, vending availability, entry allowance, and ADA compliance for each entrance and exit of each subway station in NYC. 

This dataset currently has `r nrow(nyctransit_df)` rows and `r ncol(nyctransit_df)` columns. There are `r nrow(distinct(nyctransit_df, line, station_name))` distinct stations. There are `r nrow(filter(nyctransit_df, ada == "TRUE"))` ADA-compliant stations. `r nrow(filter(nyctransit_df, vending == "NO" & entry == "TRUE"))` out of `r nrow(filter(nyctransit_df, vending == "NO"))` station entrances/exits without vending allow entrance.  

This dataset is not tidy yet because all the route variables are in their own separate columns and should be combined into just two columns, like this:

```{r combine_routes}
nyctransit_df <-
  nyctransit_df %>% 
  mutate_at(vars(route1:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    names_prefix = "route",
    values_to = "route_number") %>% 
  drop_na(route_number)
```

`r nrow(distinct(filter(nyctransit_df, route_number == "A"), line, station_name))` distinct stations serve the A train. Of these stations, `r ` are ADA-compliant.



## Problem 3

Import and clean the pols_month, snp, and unemployment datasets:

```{r clean_datasets}
polsmonth_df <-
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), "-") %>% 
  mutate_at(vars(year:day), as.numeric) %>% 
  left_join(., month_df, by = "month") %>% 
  select(!month) %>%
  relocate(year, month_name, everything()) %>% 
  rename(month = month_name) %>% 
  mutate(
    prez_gop = replace(prez_gop, prez_gop == "1", "gop"),
    prez_dem = replace(prez_dem, prez_dem == "1", "dem")
  ) %>% 
  pivot_longer(
    c(prez_gop, prez_dem),
    values_to = "president"
  ) %>% 
  filter(president != "0") %>% 
  select(!c(day, name)) %>% 
  relocate(year, month, president) %>% 
  arrange(year)

snp_df <-
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), "/") %>% 
  mutate_at(vars(month:year), as.numeric) %>% 
  left_join(., month_df, by = "month") %>% 
  select(!c(month, day)) %>%
  relocate(year, month_name, everything()) %>% 
  rename(month = month_name) %>% 
  arrange(year)

month_df <-
  mutate(month_df, month_abb = str_to_lower(month.abb))

unemployment_df <-
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month_abb",
    values_to = "unemployment_rate"
  ) %>% 
  left_join(., select(month_df, !month), by = "month_abb") %>% 
  select(!month_abb) %>%
  relocate(year, month_name, everything()) %>% 
  rename(month = month_name) %>% 
  arrange(year)
```

Now to merge all three of them together:

```{r merge_datasets}
fivethirtyeight_df <-
  left_join(polsmonth_df, snp_df, by = c("year", "month")) %>% 
  left_join(., unemployment_df, by = c("year", "month")) %>% 
  relocate(year, month, everything())
```

The final dataset is a combination of the pols-month, snp, and unemployment datasets. It contains monthly information on the number of governors, senators, and representatives in each US political party, which party the president was a member of, S&P 500 closing prices, and unemployment rates from `r min(pull(fivethirtyeight_df, year))` to `r max(pull(fivethirtyeight_df, year))`. It is `r nrow(fivethirtyeight_df)` rows by `r ncol(fivethirtyeight_df)` columns.